A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-10 shanghai.tex
FInished the proof of the new proposition in Section 9.
I'll probably edit it a bit, though, over the next few hours.

\section{Restriction to the identity component} \label{sec:restriction}

Consider the short exact sequence in the category of $\QC$-schemes
defining the component group scheme for $G$:
\begin{equation}\label{eq:pi0}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is also a $\QC$-scheme, 
Proposition~\ref{prop:pullback} implies that \eqref{eq:pi0} defines the sequence of functors
\begin{equation}\label{eq:pi1}
\begin{tikzcd}
\QC(0) \rar & \QC(\pi_0(G)) \rar & \QC(G) \rar & \QC(G^0) \rar & \QC(0)
\end{tikzcd}
\end{equation}
and therefore, after passing to isomorphism classes, the sequence of abelian groups 
\begin{equation}\label{eq:pi2}
\begin{tikzcd}
0 \rar & \QCiso{\pi_0(G)} \rar & \QCiso{G} \rar & \QCiso{G^0} \rar & 0.
\end{tikzcd}
\end{equation}
%(It is clear that $\QCiso{0}$ is trivial.)
 Note that we found the groups $\QCiso{G^0}$
and $\QCiso{\pi_0(G)}$
in Sections~\ref{sec:connected} and \ref{sec:etale}.
In this section we show that \eqref{eq:pi2} is right exact.

@@

\begin{proposition}\label{prop:restriction}
For every quasicharacter sheaf $\qcs{L}^0$ on $G^0$ there is some finite quasicharacter sheaf $\qcs{L}$ on $G$ such that $\qcs{L}\vert_{G^0} \iso \qcs{L}^0$;
in particular, the restriction functor $\QC(G)\to \QC(G^0)$ is essentially surjective.
\end{proposition}

@@

\begin{proposition}\label{prop:middleexact}
 The sequence 
 \[
  \begin{tikzcd}
  0 \rar & \QCiso{\pi_0(G)} \rar & \QCiso{G} \rar & \QCiso{G^0} \rar & 0.
  \end{tikzcd}
 \]
 is exact at $\QCiso{G}$.
\end{proposition}

\begin{proof}
Suppose $\qcs{L} \in \QC(G)$ and $\qcs{L}\vert_{G^0} \iso (\EE)_{G^0}$.
\end{proof}

\section{The sheaf--function dictionary for $\QC$-schemes}
\label{sec:snake}

>>>>

\section{Exactness of the component group sequence} \label{sec:restriction}

Consider the short exact sequence in the category of $\QC$-schemes
defining the component group scheme for $G$:
\begin{equation}\label{eq:pi0}
\begin{tikzcd}
0 \rar & G^0 \arrow{r}{\iota_0} & G \arrow{r}{\pi_0} & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is also a $\QC$-scheme, 
Proposition~\ref{prop:pullback} implies that \eqref{eq:pi0} defines the sequence of functors
\begin{equation}\label{eq:pi1}
\begin{tikzcd}
\QC(0) \rar & \QC(\pi_0(G)) \arrow{r}{\pi_0^*} & \QC(G) \arrow{r}{\iota_0^*} & \QC(G^0) \rar & \QC(0)
\end{tikzcd}
\end{equation}
and therefore, after passing to isomorphism classes, the sequence of abelian groups 
\begin{equation}\label{eq:pi2}
\begin{tikzcd}
0 \rar & \QCiso{\pi_0(G)} \arrow{r}{\pi_0^*} & \QCiso{G} \arrow{r}{\iota_0^*} & \QCiso{G^0} \rar & 0.
\end{tikzcd}
\end{equation}
%(It is clear that $\QCiso{0}$ is trivial.)
 Note that we found the groups $\QCiso{G^0}$
and $\QCiso{\pi_0(G)}$
in Sections~\ref{sec:connected} and \ref{sec:etale}.
In this section we show that \eqref{eq:pi2} is right exact.

@@

\begin{proposition}\label{prop:restriction}
For every quasicharacter sheaf $\qcs{L}^0$ on $G^0$ there is some finite quasicharacter sheaf $\qcs{L}$ on $G$ such that $\qcs{L}\vert_{G^0} \iso \qcs{L}^0$;
in particular, the restriction functor $\iota_0^* : \QC(G)\to \QC(G^0)$ is essentially surjective.
\end{proposition}

@@

\begin{proposition}\label{prop:middleexact}
 The sequence 
 \[
  \begin{tikzcd}
  0 \rar & \QCiso{\pi_0(G)} \arrow{r}{\pi_0^*} & \QCiso{G} \arrow{r}{\iota_0^*} & \QCiso{G^0} \rar & 0.
  \end{tikzcd}
 \]
 is exact at $\QCiso{G}$.
\end{proposition}

\begin{proof}
Suppose $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf on $G$ 
and $\qcs{L}\vert_{G^0} = (\EE)_{G^0}$.

For each $x\in \pi_0(\bG)$, set $\bG^x \ceq \pi_0^{-1}(x)$.
Let $g, g'$ be geometric points in the same 
geometric connected component $\bG^x$. 
Set $a = g^{-1}g'$ and note that $a$ is a geometric point in $\bG^0$. 
Let $\mu_{g,a} : \gqcs{L}_{ga} \to \gqcs{L}_g \otimes \gqcs{L}_a$ 
be the isomorphism of vector spaces obtained by restriction of 
$\mu : m^*\gqcs{L} \to \gqcs{L} \boxtimes \gqcs{L}$ to the 
geometric point $(g,a)$ on $\bG^x \times \bG^0$. 
Since $\qcs{L}\vert_{G^0} = (\EE)_{G^0}$, 
the stalk of $\gqcs{L}$ at $a$ is $\EE$. 
In this way the pair of geometric points $g, g' \in \bG^x$ 
determines an isomorphism $\varphi_{g,g'} \ceq \mu_{g,a}^{-1}$ 
from $\gqcs{L}_{g}$ to $\gqcs{L}_{g'}$.

The isomorphisms $\varphi_{g,g'}: \gqcs{L}_{g} \to \gqcs{L}_{g'}$ are canonical 
in the following sense: if $g,g'\in \bG^x$ and $h,h'\in \bG^y$ 
then it follows from \ref{QC.2} and \ref{QC.3} 
that
 \begin{equation}\label{eq:qc}
  \begin{tikzcd}[column sep=60]
   \gqcs{L}_{gh} \arrow{r}{\varphi_{gh,g'h'}} \arrow[swap]{d}{\mu_{g,h}}
  & \gqcs{L}_{g'h'} \arrow{d}{\mu_{g',h'}}
  &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} 
  &  \gqcs{L}_{\Frob{}(g)} \arrow{r}{\varphi_{\Frob{}(g),\Frob{}(g')}} \arrow[swap]{d}{\phi_{g}} & \gqcs{L}_{\Frob{}(g')} \arrow{d}{\phi_{g'}} \\
  \gqcs{L}_{g} \otimes \gqcs{L}_{h} \arrow{r}{\varphi_{g,g'}\otimes \varphi_{h,h'}}  
  & \gqcs{L}_{g'} \otimes \gqcs{L}_{h'} 
  & {}
  & \gqcs{L}_{g} \arrow{r}{\varphi_{g,g'}} & \gqcs{L}_{g'}
  \end{tikzcd}
 \end{equation}
both commute.

Now, for each $x\in \pi_0(\bG)$, pick $g_x\in \bG^x$ 
and set $\gqcs{E}_x \ceq \gqcs{L}_{g_x}$. 
Let $\phi_x : \gqcs{E}_{\Frob{}(x)} \to \gqcs{E}_x$ 
be the isomorphism of $\EE$-vector spaces obtained by composing 
$\varphi_{g_{\Frob{}(x)},\Frob{}(g_x)} : \gqcs{L}_{g_{\Frob{}(x)}} \to \gqcs{L}_{\Frob{}(g_x)}$ 
with $\phi_{g_x} : \gqcs{L}_{\Frob{}(g_x)} \to \gqcs{L}_{g_x}$. 
For each pair $x,y\in \pi_0(\bG)$ 
let $\mu_{x,y} : \gqcs{E}_{x+y}\to \gqcs{E}_x\otimes \gqcs{E}_y$ 
be the isomorphism of $\EE$-vector spaces obtained by composing
$\varphi_{g_{x+y},g_xg_y} : \gqcs{L}_{g_{x+y}} \to \gqcs{L}_{g_xg_y}$ 
with $\mu_{g_x,g_y} : \gqcs{L}_{g_xg_y} \to \gqcs{L}_{g_x}\otimes \gqcs{L}_{g_y}$.
Using \eqref{eq:qc}, it follows that \ref{qc.1}, \ref{qc.2} and \ref{qc.3} are satisfied for
\begin{enumerate}
\item
the indexed set of one-dimensional $\EE$-vector spaces  $\gqcs{E}_x$, 
as $x$ runs over $\pi_0(\bG)$,
\item the indexed set of isomorphisms
  $\mu_{x,y} : \gqcs{E}_{x+y} \xrightarrow{\iso} \gqcs{E}_{x} \otimes\gqcs{E}_{y}$,
 as $x$ and $y$ run over $\pi_0(\bG)$,
 \item the indexed set of isomorphisms $\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$,
as $x$ runs over $\pi_0(\bG)$.
\end{enumerate}
We have now defined a quasicharacter sheaf $\qcs{E}$ on $\pi_0(G)$ 
from the quasicharacter sheaf $\qcs{L}$ on $G$ 
with the property $\qcs{L}\vert_{G^0} = (\EE)_{G^0}$.

Now, pullback the quasicharacter sheaf $\qcs{E}$ along $\pi_0 : G \to \pi_0(G)$.
The resulting quasicharacter sheaf $\pi_0^* \qcs{E}$ is constant 
on geometric connected components, with stalks given by 
$(\pi_0^* \qcs{E})_g = \qcs{E}_{x}$ for all $g\in \bG^x$; 
in particular, $\left(\pi_0^* \qcs{E}\right)\vert_{\bG^x} =  (\qcs{E}_{x})_{\bG^x}$.
Likewise, the quasicharacter sheaf $\qcs{L}$ is constant 
on geometric connected components. Our choices above define isomorphisms
$\gqcs{L}\vert_{\bG^x} \to  (\gqcs{E}_{x})_{\bG^x}$, for each $x\in \pi_0(\bG)$.
The resulting isomorphism $\gqcs{L} \to \pi_0^* \gqcs{E}$ satisfies \ref{QC.4}, 
thus defining an isomorphism $\qcs{L} \to \pi_0^* \qcs{E}$.

We have now show that if $\qcs{L}$ is a quasicharacter sheaf on $G$ and $\iota_0^*\qcs{L}= (\EE)_{G^0}$ then there is some quasicharacter sheaf $\qcs{E}$ on $\pi_0(G)$ such that $\qcs{L} \iso \pi_0^* \qcs{E}$. On the other hand, $\iota_0^*(\pi_0^* \qcs{E}) \iso (\EE)_{G^0}$, by Proposition~\ref{prop:pullback}. This shows that \eqref{eq:pi2} is exact at $\QC(G)$, concluding the proof of Proposition~\ref{prop:middleexact}.
\end{proof}

\section{The sheaf--function dictionary for non-connected group schemes}
\label{sec:snake}
